---
title: "615_FinalProject_EDA"
author: "Kai"
date: "2022-12-12"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(fmsb)
library(tidyverse)
library(ggplot2)
```


```{r include=FALSE}

levels <- read.delim("/Users/dongkaiwu/Desktop/615/615_Final/MBTA_GTFS/levels.txt", ",", header = TRUE)
multi_road_trips<- read.delim("/Users/dongkaiwu/Desktop/615/615_Final/MBTA_GTFS/multi_route_trips.txt", ",", header = TRUE)
pathways<- read.delim("/Users/dongkaiwu/Desktop/615/615_Final/MBTA_GTFS/pathways.txt", ",", header = TRUE)
route_pattern<- read.delim("/Users/dongkaiwu/Desktop/615/615_Final/MBTA_GTFS/route_patterns.txt", ",", header = TRUE)
routes<- read.delim("/Users/dongkaiwu/Desktop/615/615_Final/MBTA_GTFS/routes.txt", ",", header = TRUE)
stop_times<- read.delim("/Users/dongkaiwu/Desktop/615/615_Final/MBTA_GTFS/stop_times.txt", ",", header = TRUE)
stops<- read.delim("/Users/dongkaiwu/Desktop/615/615_Final/MBTA_GTFS/stops.txt", ",", header = TRUE)
transfers<- read.delim("/Users/dongkaiwu/Desktop/615/615_Final/MBTA_GTFS/transfers.txt", ",", header = TRUE)
trips_properties<- read.delim("/Users/dongkaiwu/Desktop/615/615_Final/MBTA_GTFS/trips_properties_definitions.txt", ",", header = TRUE)
trips<- read.delim("/Users/dongkaiwu/Desktop/615/615_Final/MBTA_GTFS/trips.txt", ",", header = TRUE)

library(readr)
Travel_Times_Fall_2019 <- read_csv("Average_Rapid_Transit_Travel_Times_Fall_2019.csv")

```

# Introduction

The MBTA or the 'T' is one thing that we are most familiar with, and this EDA analysis is based on the MBTA data. The data I have used for this EDA report is from MTBA website:  https://mbta-massdot.opendata.arcgis.com/search, the data file is called Average_Rapid_Transit_Travel_Times_Fall_2019.csv. I have made three EDA analysis, covered three aspects of the Boston transit. First is the number of routes that is provided since 1901, and the second is the total number of trips been provided, last one is the travel time during the week. Now let me show you some interesting analysis about Boston transit system. 




## EDA Analysis --- Routes


```{r echo=FALSE, message=FALSE, warning=FALSE,fig.height=6,fig.width=4, fig.cap= "Number of routes for each line"}
new <- route_pattern %>% group_by(route_id) %>% summarise(total_count = n())

new2 <- filter(new, route_id %in% c("Red", "Orange", "Green-B", "Green-C", "Green-D", "Green-E"))

wider2 <- new2 %>%
  pivot_wider(names_from = route_id, values_from = total_count)

wider2 <- rbind(rep(20,5) , rep(0,5) , wider2)

colors_border=c( rgb(0.2,0.5,0.5,0.9), rgb(0.8,0.2,0.5,0.9) , rgb(0.7,0.5,0.1,0.9) )
colors_in=c( rgb(0.2,0.5,0.5,0.4), rgb(0.8,0.2,0.5,0.4) , rgb(0.7,0.5,0.1,0.4) )

radarchart( wider2  , axistype=1 , 
    #custom polygon
    pcol=colors_border , pfcol=colors_in , plwd=4 , plty=1,
    #custom the grid
    cglcol="grey", cglty=1, axislabcol="grey", caxislabels=seq(0,20,5), cglwd=0.8,
    #custom labels
    vlcex=0.8,title=paste("Number of Routes for each line"),)

```


This plot shows the number of routes provided for each line. We can find the red line has the most routes which is 14. And after that are the Green-B, Green-C, and Green-D line with a tie of 6 routes. The least is Green-E with only 4 routes. The green line is the oldest subway line in Boston, so that is why there are so many green line branchs around Boston, and add together the green line has 22 routes. 


\newpage


## EDA Analysis --- Trips


```{r echo=FALSE, message=FALSE, warning=FALSE,fig.cap= "Top 10 lines with the most number of trips"}

# Create dataset
trip_count <- trips %>% group_by(route_id) %>% summarise(total_count = n())
df<-trip_count[order(-trip_count$total_count),]
trip_count <- df[1:10,]
id = seq(1,nrow(trip_count))
total_trips <- cbind(id,trip_count)

 
# ----- This section prepare a dataframe for labels ---- #
# Get the name and the y position of each label
label_data <-  total_trips
 
# calculate the ANGLE of the labels
number_of_bar <- nrow(label_data)
angle <-  90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
 
# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
label_data$hjust<-ifelse( angle < -90, 1, 0)
 
# flip angle BY to make them readable
label_data$angle<-ifelse(angle < -90, angle+180, angle)
# ----- ------------------------------------------- ---- #
 
 
# Start the plot
p <- ggplot(total_trips, aes(x=as.factor(id), y=total_count)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  
  # This add the bars with a blue color
  geom_bar(stat="identity", fill=alpha("orange", 0.7)) +
  
  # Limits of the plot = very important. The negative value controls the size of the inner circle, the positive one is useful to add size over each bar
  ylim(-100,2600) +
  
  # Custom the theme: no axis title and no cartesian grid
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm")      # Adjust the margin to make in sort labels are not truncated!
  ) +
  
  # This makes the coordinate polar instead of cartesian.
  coord_polar(start = 0) +
  
  # Add the labels, using the label_data dataframe that we have created before
  geom_text(data=label_data, aes(x=id, y=total_count+10, label=route_id, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=3.5, angle= label_data$angle, inherit.aes = FALSE) 
 
p
```

This figure shows the top 10 lines with the most number of trips, and this kind of followed the trend on the route graph which makes sense, the red line again has the most number of trips then it followed by green lines. Those numbers represent bus lines, and among those bus lines 111 has the highest number of trips. From this graph we could see a clear comparison between bus and subways, still subway is the major public transportation is Boston.


\newpage


## EDA Analysis --- Travel Time

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.cap= "Routes Travel Time VS. Day Type"}
library(ggplot2)

ggplot(Travel_Times_Fall_2019) +
 aes(x = route_id, y = average_sec, fill = day_type) +
 geom_col() +
 scale_fill_hue(direction = 1) +
 labs(x = "Route id", y = "Average Travel Time", title = "Routes Travel Time VS. Day Type", fill = "Day Type:") +
 theme_minimal()

```

From this graph we can clearly see that all the routes travels longer during weekdays compare to weekend. And the comparison shows that the total travel time during weekday is the sum of weekend's travel time, which makes perfect sense. During weekday, people who needs to go to school or go to work all need to take subway, and that workload itself is much higher than during weekend, the time of people getting off and getting on add up to that long travel time. 


